### 2020-05-26 paper.tex
Finalize War Story Rewrite

The reports provided us diverse perspectives on various areas of software development using RxJS. One of the reporters focused on how they built code which had improved on testability over the recent versions of RxJS: ``\dots in the beginning, it was very hard to write async tests and I did not like it. I really disliked that once you started to test async code you were forced to pass a TestScheduler''. Allowing to pass the scheduler explicitly as parameter forced them to introduce code which was only necessary for testing reasons they stated further. With its current major version, RxJS 6 improved heavily on the  \mintinline{TypeScript}{TestScheduler}. The runtime environment itself can now be augmented with the scheduler which results in cleaner code.

Even though the share of non-productive, testing-related code necessary to build a mature RxJS-based application seems to have decreased today, multiple of our reporters mentioned the common practice of manually modifying production code during debugging sessions, hence confirming earlier statements from our informal interview partners. One engineer described a scenario where they suspected a problem in a complex Observable composition: Having multiple asynchronous, remote data sources, they used Observables to model the dependencies between them and implemented computations on their results as operators. On top, each data source could re-emit updated versions of previously requested information any time. Though tested thoroughly, after a week in production the first of many bugs was reported: ``Figures were changing multiple times (where I expected them to change once), other numbers were not updated at all'' they told us. The browsers debugger tools and its breakpoints did not help much since the operators were executed several multiple times and some parts of the stream were impossible to grasp, even with conditional breakpoints. ``I started to inspect the flow \dots with console.logs and later also using tags from rxjs-spy'', hence manually augmenting the code with trace logs. After the time consuming log analysis, they finally were able resolve the bug. The log statements added were removed in the aftermath, though the rxjs-spy related code was left in the Observable stream in case they might be needed again in the future. (We will elaborate on the functionality of rxjs-spy in the next section more in-depth.)

``It took quite some time to get it right and one of the most invaluable tools proved to be Stackblitz which gave us ability to quickly create  smaller working examples and iterate on them.''

We could verify previously gathered interview data with the results from the new reports: The subjects tend to manually add log statements to their code in order to understand its behavior, especially when things get more complex. Three subjects even claimed using specialized logging functionalities provided by rxjs-spy\cite{rxjsspy}. As the subjects stated, the knowledge about additional events made visible helped in situations where they needed to find problems within a graph composing multiple asynchronous data sources on multiple levels. Without it, they would have been blind to what actually happened within the system.

In addition to more sophisticated logging, three subjects reported that they built a habit of decomposing complex graphs into smaller, simpler bits and running those in a dedicated environment like Rx Visualizer\cite{rxviz} or StackBlitz\footnote{\url{https://stackblitz.com/}}.

>>>>

The reports provided us diverse perspectives on various areas of software development using RxJS. One of the reporters focused on how they built code which had improved on testability over the recent versions of RxJS: ``[\dots] in the beginning, it was very hard to write async tests and I did not like it. I really disliked that once you started to test async code you were forced to pass a TestScheduler''. Allowing to pass the scheduler explicitly as parameter forced them to introduce code which was only necessary for testing reasons they stated further. With its current major version, RxJS 6 improved heavily on the  \mintinline{TypeScript}{TestScheduler}. The runtime environment itself can now be augmented with the scheduler which results in cleaner code.

Even though the share of non-productive, testing-related code necessary to build a mature RxJS-based application seems to have decreased today, multiple of our reporters mentioned the common practice of manually modifying production code during debugging sessions, hence confirming earlier statements from our informal interview partners. One engineer described a specific scenario where they suspected a problem within a complex Observable composition: Having multiple asynchronous, remote data sources, they used Observables to model the dependencies between them and implemented computations on their results as operators. On top, each data source could re-emit updated versions of previously requested information any time. Though tested thoroughly, after a week in production the first of many bugs was reported: ``Figures were changing multiple times (where I expected them to change once), other numbers were not updated at all'' they told us. The browsers debugger tools and its breakpoints did not help much since the operators were executed several multiple times and some parts of the stream were impossible to grasp, even with conditional breakpoints. ``I started to inspect the flow [\dots] with console.logs and later also using tags from rxjs-spy'', hence manually augmenting the code with trace logs. After the time consuming log analysis, they finally were able resolve the bug. The log statements added were removed in the aftermath, though the rxjs-spy related code was left in the Observable stream in case they might be needed again in the future. (We will elaborate on the functionality of rxjs-spy in Section~\ref{sec:discussion} more in-depth.)

In a related war story, another engineer disclosed similar practices using  external tools: ``It took quite some time to get it right and one of the most invaluable tools proved to be Stackblitz which gave us ability to quickly create smaller working examples and iterate on them.'' Stackblitz\footnote{\url{https://stackblitz.com/}} is a full JavaScript development environment available online. They used this sandboxed setting to run, debug and iterate on selected pieces of a larger Observable stream. Even though the final result had to be integrated back into the actual application, the extra effort was worth the result the engineer concluded in their report.
